\LARGE \bf On the use of vehicle categorization and context information for messages forwarding in VANETs

### 2016-09-17 paperITSC.tex
USING ONTOLOGY FOR VEHICLES CATEGORIZATION text

\begin{table}[ht]
    \center
    %\renewcommand{\arraystretch}{1.2}
    \caption{Simulation Parameters} \label{simulation-param}
    \begin{tabular}{|c|c|}
        \hline
        Packet format                  & WAVE short message \\ \hline
        Status beacon Interval         & 1 second           \\ \hline
        Validity of status beacons     & 3 seconds          \\ \hline
        Message creation interval      & 15 seconds         \\ \hline
        Beacon Length                  & 512 KB             \\ \hline
        Destination Position           & (1500, 1500)       \\ \hline
        Transmission range of vehicles & 250 meters         \\ \hline
        Simulation time                & 600 seconds        \\ \hline
        Maximum number of hops         & 10 hops            \\ \hline
    \end{tabular}
\end{table}

In \emph{MFCV} we assume that each vehicle is able to collect the required contextual information about itself and store information received from its neighbors. Furthermore, the contextual information shared among the vehicles is reliable and the vehicles collaborate with the forwarding process, since security is not the focus of this work. In the following subsections, we define the models used for the implementation.

@@

We developed a mobility scenario that consists of a grid of 9 km$^2$ (urban area) using the \emph{netgenerate} tool from SUMO. In order to evaluate the influence of the different categories of vehicles in the forwarding decision, 10 vehicles were classified as taxis, and the remaining as private vehicles. We used \emph{randomTrips}, SUMO for generating routes, in which the taxis accomplished 15 km while the private vehicles only 2 km. Due to this, the taxi will take maximum simulation time (i.e., 10 minutes) to complete its route, whereas the private vehicles leave the scenario in approximately 2 minutes. This way, the taxis can increase the successful delivery rate of messages as compared to private vehicles.

Initially, 10 taxi and 40 private vehicles enter in the scenario. After 2 minutes, additional 40 private vehicles enter in the scenario. After an interval of 2 minutes, 40 new private vehicles enter the scenario. This ensure that at any time the ratio of private vehicles to taxi will be 4 to 1. A total of 210 vehicles (200 private vehicles and 10 taxis) participate in the experiment.

@@

The first phase of \emph{MFCV} is to discovery the neighborhood of a vehicle, and creates a \emph{neighbors table} (hash table). The table contains information about the neighbors ID, its current and a previous geographical position (5 seconds ago), categories (type and use), speed and a timestamp. Each vehicle broadcasts this information, with an interval of 1 second, in a message called \emph{status beacon}. When a vehicle receives a \emph{status beacon}, it can either store it as a new entry in the table or update an entry, if it is already exists. This information remains valid for 3 seconds in the vehicle buffer. This information help the vehicles to make more accurate decisions in the \emph{Forwarding Phase}.

@@

In the second phase, the messages are forwarded, using \emph{neighbors table} information, to improve the forwarding decisions. Random vehicles are chosen to create the messages during the simulation. When a vehicle enters the scenario, it may create a new message and a schedule to try to send it according to its \emph{RateTimeToSend} value. The content of this message contains 4 IDs (i.e, source, temporarySender, temporaryReceiver, and destination), the hopCount (number of hops to reach the destination), the TTL, the message by itself and timestamp. The vehicle will try to send the message to the neighbor with the best probability, based in the information in the neighbors table, to delivery the message. If the destination is one of the neighbor, the message is delivery to it. The message has a maximum allowed time (TTL - Time To Live) in the simulation. The value of TTL is 60 or 120 seconds. Once the TTL is expired, the message is discarded from the buffer. Moreover, if a node receives a message and its buffer is full then the oldest message is discarded.

The decision making process was divided into three different schemes. In the first scheme the node with the shortest distance to the destination was chosen to carry the message. The distance and speed of the nodes were used in the second approach. In the third scheme, another information was exploited, i.e., the categories \emph{taxi} and \emph{private vehicles} and its respective mobility patterns. In all experiment, we used a path history of each vehicle, that contains its current and 5 seconds old position, as a way to know if the vehicle is moving towards the destination, hence being a good option to carry the message.

Due to wide and flexible mobility, the \emph{taxi} has a higher chance to deliver the message successfully as compared to the \emph{private vehicles}. To prevent the \emph{taxis} to be overloaded, \emph{taxis} had a 80\% chance of selection, in the decision process. Additionally, we developed a mechanism called \emph{RateTimeToSend} that is a value, linked with the vehicle speed, it ensure that vehicles with low speed (under 10 m/s) do not flood the network. In case of traffic jams it controls the amount of messages forwarded.

Initially, its value is 2.5 seconds for all vehicles. After 2.5 seconds the vehicle tries to forward the first message in the buffer, and schedule to forward the next message in 2.4 or 2.6 seconds according to its mobility. The value increases 0.1 s, if the vehicle does not move 10 meters in 1 second, otherwise, the value decreases 0.1 s. The value is always between 0.1 and 5 seconds. When the vehicle run its whole vector it will check if have passed 5 seconds since it send the first message, if so, it will start to send the messages from the beginning of the buffer (just as before), if it is not the case it will wait for 5 seconds to start from the beginning.

@@

In our simulation scenario, we installed a single Road-Side Unit (RSU), in the middle of the scenario. This RSU is the destination point of all messages. The vehicles (both private vehicles  and taxis), enters the scenario, at point defined by its routes. We run the simulation for a total of 10 minutes. Furthermore, the buffer was capable of handling 50 messages at a time.

@@

Depending on the experiment, the number of vehicles that generate the message, every 15 seconds, can be either 1 or 2. To guarantee that the message would stay in the scenario its TTL (i.e., 60, 120 or 150s according to the experiment) the messages will be created until 450 seconds of simulation time. Hence, at the end of the simulation a total of 30 or 60 messages are created.

>>>>

\begin{table}[ht]
    \center
    %\renewcommand{\arraystretch}{1.2}
    \caption{Simulation Parameters} \label{simulation-param}
    \begin{tabular}{|c|c|}
        \hline
        Packet format                  & WAVE short message \\ \hline
        Status beacon Interval         & 1 second           \\ \hline
        Validity of status beacons     & 3 seconds          \\ \hline
        Message creation interval      & 15 seconds         \\ \hline
        Beacon Length                  & 512 KB             \\ \hline
        Destination Position           & (1500, 1500)       \\ \hline
        Transmission ranges of vehicles & 250 meters         \\ \hline
        Simulation time                & 600 seconds        \\ \hline
        Maximum number of hops         & 10 hops            \\ \hline
    \end{tabular}
\end{table}

In \emph{MFCV} we assume that each vehicle is able to collect the required contextual information about itself and store information received from its neighbors. Furthermore, the contextual information shared among the vehicles is reliable and the vehicles collaborate with the forwarding process since security is not the focus of this work. In the following subsections, we define the models used for the implementation.

@@

We developed a mobility scenario that consists of a grid of 9 km$^2$ (urban area) using the \emph{netgenerate} tool from SUMO. In order to evaluate the influence of the different categories of vehicles in the forwarding decision, 10 vehicles were classified as taxis and the remaining as private vehicles. We used \emph{randomTrips}, SUMO for generating routes, in which the taxis accomplished 15 km while the private vehicles only 2 km. Due to this, the taxi will take maximum simulation time (i.e., 10 minutes) to complete its route, whereas the private vehicles leave the scenario in approximately 2 minutes. This way, the taxis can increase the successful delivery rate of messages as compared to private vehicles.

Initially, 10 taxi and 40 private vehicles enter in the scenario. After 2 minutes, additional 40 private vehicles enter in the scenario. After an interval of 2 minutes, 40 new private vehicles enter the scenario. This ensures that at any time the ratio of private vehicles to taxis will be 4 to 1. A total of 210 vehicles (200 private vehicles and 10 taxis) participate in the experiment.

@@

The first phase of \emph{MFCV} is to discover the neighborhood of a vehicle and creates a \emph{neighbors' table} (hash table). The table contains information about the neighbors' ID, its current and a previous geographical position (5 seconds ago), categories (type and use), speed and a timestamp. Each vehicle broadcasts this information, with an interval of 1 second, in a message called \emph{status beacon}. When a vehicle receives a \emph{status beacon}, it can either store it as a new entry in the table or update an entry, if it already exists. This information remains valid for 3 seconds in the vehicle buffer. This information helps the vehicles to make more accurate decisions in the \emph{Forwarding Phase}.

@@

In the second phase, the messages are forwarded, using \emph{neighbors' table} information, to improve the forwarding decisions. Random vehicles are chosen to create the messages during the simulation. When a vehicle enters the scenario, it may create a new message and a schedule to try to send it according to its \emph{RateTimeToSend} value. The content of this message contains 4 IDs (i.e, source, temporarySender, temporaryReceiver, and destination), the hopCount (number of hops to reach the destination), the TTL, the message by itself and timestamp. The vehicle will try to send the message to the neighbor with the best probability, based on the information in the neighbors' table, to deliver the message. If the destination is one of the neighbors, the message is delivery to it. The message has a maximum allowed time (TTL - Time To Live) in the simulation. The value of TTL is 60 or 120 seconds. Once the TTL is expired, the message is discarded from the buffer. Moreover, if a node receives a message and its buffer is full then the oldest message is discarded.

The decision-making process was divided into three different schemes. In the first scheme, the node with the shortest distance to the destination was chosen to carry the message. The distance and speed of the nodes were used in the second approach. In the third scheme, another information was exploited, i.e., the categories \emph{taxi} and \emph{private vehicles} and its respective mobility patterns. In all experiment, we used a path history of each vehicle, that contains its current and 5 seconds old position, as a way to know if the vehicle is moving towards the destination, hence being a good option to carry the message.

Due to wide and flexible mobility, the \emph{taxi} has a higher chance to deliver the message successfully as compared to the \emph{private vehicles}. To prevent the \emph{taxis} to be overloaded, \emph{taxis} had a 80\% chance of selection, in the decision process. Additionally, we developed a mechanism called \emph{RateTimeToSend} that is a value, linked with the vehicle speed, it ensures that vehicles with low speed (under 10 m/s) do not flood the network. In the case of traffic jams it controls a number of messages forwarded.

Initially, its value is 2.5 seconds for all vehicles. After 2.5 seconds the vehicle tries to forward the first message in the buffer, and schedule to forward the next message in 2.4 or 2.6 seconds according to its mobility. The value increases 0.1 seconds if the vehicle does not move 10 meters in 1 second, otherwise, the value decreases 0.1 s. The value is always between 0.1 and 5 seconds. When the vehicle run its whole vector it will check if have passed 5 seconds since it sends the first message, if so, it will start to send the messages from the beginning of the buffer (just as before), if it is not the case it will wait for 5 seconds to start from the beginning.

@@

In our simulation scenario, we installed a single Road-Side Unit (RSU), in the middle of the scenario. This RSU is the destination point of all messages. The vehicles (both private vehicles  and taxis), enters the scenario, at the point defined by its routes. We run the simulation for a total of 10 minutes. Furthermore, the buffer was capable of handling 50 messages at a time.

@@

Depending on the experiment, the number of vehicles that generate the message, every 15 seconds, can be either 1 or 2. To guarantee that the message would stay in the scenario its TTL (i.e., 60, 120 or 150s according to the experiment) the messages will be created until 450 seconds of simulation time. Hence, at the end of the simulation, a total of 30 or 60 messages are created.
